---
title: "SIZE_change"
author: "Elyza Pilipaityte"
date: "2024-03-27"
output: html_document
---

## clear memory
```{r}
rm(list=ls())
```

Loading libraries

```{r}
library(tidyverse)
library(car)
# library(dplyr)
library(ggplot2)
library(lme4)
# install.packages("sjPlot")
library(sjPlot)
library(GGally)
library(DHARMa)

```

Loading packages

```{r}
temp <- read_csv(file = "temp.csv")
size_age_data <- read_csv(file = "size_at_age.csv")

```


# Bream

```{r}
bream <- size_age_data %>% filter (spp_name_lt=="Karsis")

colSums(is.na(bream))

# Creating new column (year1) by subtracting fish age from the year when it was caught. It will show first fish year - the year the fish was born.
bream$year1 <- bream$year - bream$age

# Renaming column to year1 in temp dataset in order to combine it with perch data
names(temp)[names(temp) == "year"] <- "year1"

# Merging perch and temp dataset by year1 - just adding temp column which will show water temperature for first year of fish life
bream <- merge(bream, temp, by = "year1", all.x = TRUE)

# Creating new column for average temperature during all years of fish life
average_temp <- numeric(nrow(bream))

for (i in 1:nrow(bream)) {
  years_subset <- temp$temp[temp$year1 >= bream[i, "year1"] & temp$year1 <= bream[i, "year"]]
  average_temp[i] <- mean(years_subset, na.rm = TRUE)
}

bream$average_temp <- average_temp

# renaming columns to make it more clear
names(bream)[names(bream) == "temp"] <- "temp_year1"
names(bream)[names(bream) == "average_temp"] <- "temp_life"
```

Tables to see number of observations and main statiscis

```{r}
tableb1<- bream %>%
  group_by(year, method) %>%
  summarise(
    median_length = median(std_length),
    median_age = median(age), count_observation = n())

# scales and otholits data for age groups
tableb2<- bream %>%
  group_by(age, method) %>%
  summarise(count = n(), .groups = 'drop')  %>%
  pivot_wider(names_from = method, values_from = count, values_fill = 0)

```

Data exploration

```{r}
colSums(is.na(bream))
table(bream$age)

# removing ages which do not have both methods
bream1 <- bream %>% filter (age > 2 & age <= 16 & age != 5) %>% drop_na (season) %>% filter (std_length<55) 

# making one group of ages between 11 to 16 years
bream1 <- bream1 %>%
  mutate(age = ifelse(age > 11, 12, age))

table(bream1$year)

bream1 <- bream1 %>% mutate(yearn = year - 1948)
table(bream1$method)

bream1$fage <- as.factor(bream1$age)
```


1. Outliers in response and independent variables

```{r}
ggplot(data = bream1) + 
  geom_point(mapping = aes(x = age, y = std_length, color=as.factor(year)), position = "jitter", size = 2)
# no outliers
```

2. Normality and homogeneity of the response variable

```{r}
# Frequency plots
par(mfrow = c(1,2), mar = c(4,5,2,3), cex.lab = 1)
hist(bream1$std_length,  xlab = "Length (cm)",
     col = "lightblue", border = "black", main = "")
hist((bream1$age),  xlab = "Age", 
     col = "red", border = "black",  main = "")

```

3. Balance of categorical variables

```{r}
table(bream1$season)
```

4. Multicollinearity among covariates

```{r}
Coll_b <- c("std_length", "age", "temp_life", "method", "season")
# Obtain summary using the ggpairs command from the GGally library
ggpairs(bream1[ ,Coll_b], ggplot2::aes(alpha = 0.8, colour = season))
# No obvious collinearity

#Calculate Variance Inflation Factor (VIF)
round(vif(lm(std_length ~ age + temp_life + method + season,
                     data = bream1)),2)

```

5. Relationships among response and independent variables

```{r}
# Interaction? Plot data
ggplot(bream1, aes(x = age, y = (std_length), color = season)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("Age") + ylab("Length (cm)") +
  facet_wrap(~method)

ggplot(bream1, aes(x = temp_life, y = (std_length), color = season)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("temperature1") + ylab("Length (cm)") +
  facet_wrap(~method)
```

Finding the best model
```{r}

P1 <- lmer(std_length ~ age + temp_life +  age * method + (1|season), data = bream1, REML = F)
P2b <- lmer(std_length ~ age * temp_life +  age * method + (1|season), data = bream1, REML = F)
anova(P1, P2b)

P3 <- lmer(std_length ~ fage + temp_life +  fage * method + (1|season), data = bream1, REML = F)
P4b <- lmer(std_length ~ fage * temp_life +  fage * method + (1|season), data = bream1, REML = F)
anova(P3, P4b)
```

```{r}
plot_model(P2b, type = "pred", 
                   terms = c("temp_life", "age"),
                   show.data = TRUE,
                   pred.type = c("fe"),
                   title = "",
                   show.legend = TRUE,
                   jitter = 0.1,
                   axis.title = c("Average fish life temperature", "Standard length (cm)"),
                   show.values = FALSE) +
   ylim(0, NA) +
  theme_bw() +  # Set plot theme to black and white
  theme(panel.background = element_rect(fill = "white"),  # Set background color to white
        axis.text = element_text(size = 14),  # Increase size of axis text
        axis.title = element_text(size = 16),  # Increase size of axis titles
        panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank())  # Remove minor grid lines

```

Using age as factor. Model with different slopes for different age group.
```{r}
plot_model(P4b,
                   type = "pred", 
                   terms = c("temp_life", "fage"),
                   show.data = TRUE,
                   pred.type = c("fe"),
                   title = "",
                   show.legend = TRUE,
                   jitter = 0.1,
                   axis.title = c("Average fish life temperature", "Standard length (cm)"),
                   show.values = FALSE) +
  theme_bw() +  # Set plot theme to black and white
  theme(panel.background = element_rect(fill = "white"),  # Set background color to white
        axis.text = element_text(size = 14),  # Increase size of axis text
        axis.title = element_text(size = 16),  # Increase size of axis titles
        panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank())  # Remove minor grid lines

```

```{r}
# Summarise neatly in a table
tab_model(P2b,
          show.zeroinf = F,
             dv.labels = c("Bream"),
           string.pred = "Coefficient",
             string.ci = "Conf. Int (95%)",
              string.p = "P-value",
               p.style = c("numeric"),
                emph.p = FALSE,
             transform = NULL)
```


Residuals
```{r}
plot(P2b)
resid <- residuals(P2b)
hist(resid, breaks = 30, main = "Histogram of Residuals")
qqnorm(resid)
qqline(resid, col = "red")

```

```{r}
# Influential observations in the model? 
# Measure using Cook's distance
par(mfrow = c(1, 1))
plot(cooks.distance(P2b), 
     type = "h",
     xlab = "Observation", 
     ylab = "Cook's distance",
     ylim = c(-0.1, 1.2))
abline(h=1, lty = 2)
abline(h=0, lty = 2)
# Not good
```


# Roach

```{r}
roach <- size_age_data %>% filter (spp_name_lt=="Kuoja")

colSums(is.na(roach))

# Creating new column (year1) by subtracting fish age from the year when it was caught. It will show first fish year - the year the fish was born.
roach$year1 <- roach$year - roach$age

# Renaming column to year1 in temp dataset in order to combine it with perch data
names(temp)[names(temp) == "year"] <- "year1"

# Merging perch and temp dataset by year1 - just adding temp column which will show water temperature for first year of fish life
roach <- merge(roach, temp, by = "year1", all.x = TRUE)

# Creating new column for average temperature during all years of fish life
average_temp <- numeric(nrow(roach))

for (i in 1:nrow(roach)) {
  years_subset <- temp$temp[temp$year1 >= roach[i, "year1"] & temp$year1 <= roach[i, "year"]]
  average_temp[i] <- mean(years_subset, na.rm = TRUE)
}

roach$average_temp <- average_temp

# perch <- perch %>% drop_na(temp_year1)

# renaming columns to make it more clear
names(roach)[names(roach) == "temp"] <- "temp_year1"
names(roach)[names(roach) == "average_temp"] <- "temp_life"
```

Tables to see number of observations and main statiscis

```{r}
tabler1 <- roach %>%
  group_by(year, method) %>%
  summarise(
    median_length = median(std_length),
    median_age = median(age), count_observation = n())

# scales and otholits data for age groups
tabler2<- roach %>%
  group_by(age, method) %>%
  summarise(count = n(), .groups = 'drop')  %>%
  pivot_wider(names_from = method, values_from = count, values_fill = 0)
```

Data exploration

```{r}
colSums(is.na(roach))
table(roach$age)
# removing ages which do not have both methods
roach1 <- roach %>% filter (age > 0 & age <= 14) %>% drop_na(season)

# making one group of ages between 11 to 16 years
roach1 <- roach1 %>%
  mutate(age = ifelse(age > 8, 9, age))

table(roach1$year)

roach1 <- roach1 %>% mutate(yearn = year - 1949) %>% filter(std_length<55)
table(roach1$yearn)

roach1$fage <- as.factor(roach1$age)
```

1. Outliers in response and independent variables

```{r}
ggplot(data = roach1) + 
  geom_point(mapping = aes(x = age, y = std_length, color=as.factor(year)), position = "jitter", size = 2)
# no obvious outliers
```

2. Normality and homogeneity of the response variable

```{r}
# Frequency plots
par(mfrow = c(1,2), mar = c(4,5,2,3), cex.lab = 1)
hist(roach1$std_length,  xlab = "Length (cm)",
     col = "lightblue", border = "black", main = "")
hist((roach1$age),  xlab = "Age", 
     col = "red", border = "black",  main = "")

```

3. Balance of categorical variables

```{r}
table(roach1$season)
```

4. Multicollinearity among covariates

```{r}
Coll_b <- c("std_length", "age", "temp_life", "method", "season")
# Obtain summary using the ggpairs command from the GGally library
ggpairs(roach1[ ,Coll_b], ggplot2::aes(alpha = 0.8, colour = season))
# No obvious collinearity

#Calculate Variance Inflation Factor (VIF)
round(vif(lm(std_length ~ age + temp_life + method + season,
                     data = roach1)),2)

```

5. Relationships among response and independent variables

```{r}
# Interaction? Plot data
ggplot(roach1, aes(x = age, y = (std_length), color = season)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("Age") + ylab("Length (cm)") +
  facet_wrap(~method)

ggplot(roach1, aes(x = temp_year1, y = (std_length), color = season)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("temperature1") + ylab("Length (cm)") +
  facet_wrap(~method)
```

Finding the best model
```{r}

P1 <- lmer(std_length ~ age + temp_life + age * method  + (1|season), data = roach1, REML = F)
P2r <- lmer(std_length ~ age * temp_life + age * method + (1|season), data = roach1, REML = F)
anova(P1, P2r)

P3 <- lmer(std_length ~ fage + temp_life + fage * method  + (1|season), data = roach1, REML = F)
P4r <- lmer(std_length ~ fage * temp_life + fage * method + (1|season), data = roach1, REML = F)
anova(P3, P4r)
```


```{r}
plot_model(P2r,
                   type = "pred", 
                   terms = c("temp_life", "age"),
                   show.data = TRUE,
                   pred.type = c("fe"),
                   title = "",
                   show.legend = TRUE,
                   jitter = 0.1,
                   axis.title = c("Average fish life temperature", "Standard length (cm)"),
                   show.values = FALSE) +
  ylim(0, NA) +
  theme_bw() +  # Set plot theme to black and white
  theme(panel.background = element_rect(fill = "white"),  # Set background color to white
        axis.text = element_text(size = 14),  # Increase size of axis text
        axis.title = element_text(size = 16),  # Increase size of axis titles
        panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank())  # Remove minor grid lines

```

```{r}
plot_model(P4r,
                   type = "pred", 
                   terms = c("temp_life", "fage"),
                   show.data = TRUE,
                   pred.type = c("fe"),
                   title = "",
                   show.legend = TRUE,
                   jitter = 0.1,
                   axis.title = c("Average fish life temperature", "Standard length (cm)"),
                   show.values = FALSE) +
  theme_bw() +  # Set plot theme to black and white
  theme(panel.background = element_rect(fill = "white"),  # Set background color to white
        axis.text = element_text(size = 14),  # Increase size of axis text
        axis.title = element_text(size = 16),  # Increase size of axis titles
        panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank())  # Remove minor grid lines

```

```{r}
# Summarise neatly in a table
tab_model(P2r,
          show.zeroinf = F,
             dv.labels = c("Roach"),
           string.pred = "Coefficient",
             string.ci = "Conf. Int (95%)",
              string.p = "P-value",
               p.style = c("numeric"),
                emph.p = FALSE,
             transform = NULL)
```


Residuals
```{r}

plot(P2r)
resid <- residuals(P2r)
hist(resid, breaks = 30, main = "Histogram of Residuals")
qqnorm(resid)
qqline(resid, col = "red")

```

```{r}
# Influential observations in the model? 
# Measure using Cook's distance
par(mfrow = c(1, 1))
plot(cooks.distance(P2r), 
     type = "h",
     xlab = "Observation", 
     ylab = "Cook's distance",
     ylim = c(-0.1, 1.2))
abline(h=1, lty = 2)
abline(h=0, lty = 2)
# All below 1 - that is fine
```


# Perch 

```{r}
perch <- size_age_data %>% filter (spp_name_lt=="Eserys")

colSums(is.na(perch))

# Creating new column (year1) by subtracting fish age from the year when it was caught. It will show first fish year - the year the fish was born.
perch$year1 <- perch$year - perch$age

# Renaming column to year1 in temp dataset in order to combine it with perch data
names(temp)[names(temp) == "year"] <- "year1"

# Merging perch and temp dataset by year1 - just adding temp column which will show water temperature for first year of fish life
perch <- merge(perch, temp, by = "year1", all.x = TRUE)

# Creating new column for average temperature during all years of fish life
average_temp <- numeric(nrow(perch))

for (i in 1:nrow(perch)) {
  years_subset <- temp$temp[temp$year1 >= perch[i, "year1"] & temp$year1 <= perch[i, "year"]]
  average_temp[i] <- mean(years_subset, na.rm = TRUE)
}

perch$average_temp <- average_temp

# perch <- perch %>% drop_na(temp_year1)

# renaming columns to make it more clear
names(perch)[names(perch) == "temp"] <- "temp_year1"
names(perch)[names(perch) == "average_temp"] <- "temp_life"

```

Tables to see number of observations and main statiscis

```{r}
tablep1 <- perch %>%
  group_by(year, method) %>%
  summarise(
    median_length = median(std_length),
    median_age = median(age), count_observation = n())

# scales and otholits data for age groups
tablep2<- perch %>%
  group_by(age, method) %>%
  summarise(count = n(), .groups = 'drop')  %>%
  pivot_wider(names_from = method, values_from = count, values_fill = 0)
```

Data exploration

```{r}
colSums(is.na(perch))
table(perch$age)
# removing ages which do not have both methods
perch1 <- perch %>% filter (age > 1 & age < 9) %>%  drop_na(season)
table(perch1$year)

perch1 <- perch1 %>% mutate(yearn = year - 1948)
table(perch1$yearn)

perch1$fage <- as.factor(perch1$age)
```


1. Outliers in response and independent variables

```{r}
ggplot(data = perch1) + 
  geom_point(mapping = aes(x = age, y = total_length, color=as.factor(year)), position = "jitter", size = 2)
# no outliers
```

2. Normality and homogeneity of the response variable

```{r}
# Frequency plots
par(mfrow = c(1,2), mar = c(4,5,2,3), cex.lab = 1)
hist(perch1$total_length,  xlab = "Length (cm)",
     col = "lightblue", border = "black", main = "")
hist((perch1$age),  xlab = "Age", 
     col = "red", border = "black",  main = "")

```

3. Balance of categorical variables

```{r}
table(perch1$season)
```

4. Multicollinearity among covariates

```{r}
Coll_b <- c("total_length", "age", "temp_year1", "method", "season")
# Obtain summary using the ggpairs command from the GGally library
ggpairs(perch1[ ,Coll_b], ggplot2::aes(alpha = 0.8, colour = season))
# No obvious collinearity

#Calculate Variance Inflation Factor (VIF)
round(vif(lm(total_length ~ age + temp_life + method + season,
                     data = perch1)),2)

```

5. Relationships among response and independent variables

```{r}
# Interaction? Plot data
ggplot(perch1, aes(x = age, y = (total_length), color = season)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("Age") + ylab("Length (cm)") +
  facet_wrap(~method)

ggplot(perch1, aes(x = temp_life, y = (total_length), color = season)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("temperature") + ylab("Length (cm)") +
  facet_wrap(~method)
```

Finding the best model
```{r}
P1 <- lmer(total_length ~ age + temp_life + age * method + (1|season), data = perch1, REML = F)
P2p <- lmer(total_length ~ age * temp_life + age * method + (1|season), data = perch1, REML = F)
anova(P1, P2p)

P3 <- lmer(total_length ~ fage + temp_life + fage * method + (1|season), data = perch1, REML = F)
P4p <- lmer(total_length ~ fage * temp_life + fage * method + (1|season), data = perch1, REML = F)
anova(P3, P4p)
```

```{r}
plot_model(P2p,
                   type = "pred", 
                   terms = c("temp_life", "age"),
                   show.data = TRUE,
                   pred.type = c("fe"),
                   title = "",
                   show.legend = TRUE,
                   jitter = 0.1,
                   axis.title = c("Average fish life temperature", "Total length (cm)"),
                   show.values = FALSE) +
  theme_bw() +  # Set plot theme to black and white
  theme(panel.background = element_rect(fill = "white"),  # Set background color to white
        axis.text = element_text(size = 14),  # Increase size of axis text
        axis.title = element_text(size = 16),  # Increase size of axis titles
        panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank())

```

Using age as factor. Model with different slopes for different age group.
```{r}
plot_model(P4p,
           type = "pred", 
           terms = c("temp_life", "fage"),
           show.data = TRUE,
           pred.type = c("fe"),
           title = "",
           show.legend = TRUE,
           jitter = 0.1,
           axis.title = c("Average fish life temperature", "Total length (cm)"),
           show.values = FALSE) +
  scale_x_continuous(breaks = c(8, 9, 10)) +  # Set custom breaks for the x-axis
  theme_bw() +  # Set plot theme to black and white
  theme(panel.background = element_rect(fill = "white"),  # Set background color to white
        axis.text = element_text(size = 14),  # Increase size of axis text
        axis.title = element_text(size = 16),  # Increase size of axis titles
        panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank())    # Remove minor grid lines

```


```{r}
# Summarise neatly in a table
tab_model(P2p,
          show.zeroinf = F,
             dv.labels = c("Perch"),
           string.pred = "Coefficient",
             string.ci = "Conf. Int (95%)",
              string.p = "P-value",
               p.style = c("numeric"),
                emph.p = FALSE,
             transform = NULL)
```

The assumptions, for a linear mixed effects model:
• The explanatory variables are related linearly to the response.
• The errors have constant variance.
• The errors are independent.
• The errors are normally distributed.


Residuals
```{r}
plot(P2p)
resid <- residuals(P2p)
hist(resid, breaks = 30, main = "Histogram of Residuals")
qqnorm(resid)
qqline(resid, col = "red")

```

```{r}
# Influential observations in the model? 
# Measure using Cook's distance
par(mfrow = c(1, 1))
plot(cooks.distance(P2p), 
     type = "h",
     xlab = "Observation", 
     ylab = "Cook's distance",
     ylim = c(-0.1, 1.2))
abline(h=1, lty = 2)
abline(h=0, lty = 2)
# All below 1 - that is fine
```


# Pikeperch

```{r}
sander <- size_age_data %>% filter (spp_name_lt=="Starkis")

colSums(is.na(sander))

# Creating new column (year1) by subtracting fish age from the year when it was caught. It will show first fish year - the year the fish was born.
sander$year1 <- sander$year - sander$age

# Renaming column to year1 in temp dataset in order to combine it with perch data
names(temp)[names(temp) == "year"] <- "year1"

# Merging perch and temp dataset by year1 - just adding temp column which will show water temperature for first year of fish life
sander <- merge(sander, temp, by = "year1", all.x = TRUE)

# Creating new column for average temperature during all years of fish life
average_temp <- numeric(nrow(sander))

for (i in 1:nrow(sander)) {
  years_subset <- temp$temp[temp$year1 >= sander[i, "year1"] & temp$year1 <= sander[i, "year"]]
  average_temp[i] <- mean(years_subset, na.rm = TRUE)
}

sander$average_temp <- average_temp

# renaming columns to make it more clear
names(sander)[names(sander) == "temp"] <- "temp_year1"
names(sander)[names(sander) == "average_temp"] <- "temp_life"
```

Tables to see number of observations and main statiscis

```{r}
tables1 <- sander %>%
  group_by(year, method) %>%
  summarise(
    median_length = median(std_length),
    median_age = median(age), count_observation = n())

# scales and otholits data for age groups
tables2 <- sander %>%
  group_by(age, method) %>%
  summarise(count = n(), .groups = 'drop')  %>%
  pivot_wider(names_from = method, values_from = count, values_fill = 0)
```

Data exploration

```{r}
table(sander$age)

# including only those ages with both aging methods which are up to 6 years
sander1 <- sander %>% filter (age <= 5) %>% drop_na(season)
table(sander1$year)

colSums(is.na(sander1))

sander1$fage <- as.factor(sander1$age)
```

1. Outliers in response and independent variables

```{r}
ggplot(data = sander1) + 
  geom_point(mapping = aes(x = age, y = std_length, color=as.factor(year)), position = "jitter", size = 2)

```

2. Normality and homogeneity of the response variable

```{r}
# Frequency plots
par(mfrow = c(1,2), mar = c(4,5,2,3), cex.lab = 1)
hist(sander1$std_length,  xlab = "Length (cm)",
     col = "lightblue", border = "black", main = "")
hist((sander1$age),  xlab = "Age", 
     col = "red", border = "black",  main = "")

# Data looks normally distributed
```

3. Balance of categorical variables

```{r}
table(sander1$season)
```

4. Multicollinearity among covariates

```{r}
Coll_b <- c("std_length", "age", "temp_life", "method", "season")
# Obtain summary using the ggpairs command from the GGally library
ggpairs(sander1[ ,Coll_b], ggplot2::aes(alpha = 0.8, colour = season))
# No obvious collinearity

#Calculate Variance Inflation Factor (VIF)
round(vif(lm(std_length ~ age + temp_life + method + season,
                     data = sander1)),2)

```

5. Relationships among response and independent variables

```{r}
# Interaction? Plot data
ggplot(sander1, aes(x = age, y = (std_length), color = season)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("Age") + ylab("Length (cm)") +
  facet_wrap(~method)

ggplot(sander1, aes(x = temp_life, y = (std_length), color = season)) +
  geom_point(shape = 16, size = 2, alpha = 0.7, position = "jitter") +
  geom_smooth(method = 'lm', colour = 'red', se = FALSE, size = 1.5) +
  xlab("temperature1") + ylab("Length (cm)") +
  facet_wrap(~method)
```


Finding the best model
```{r}

P1 <- lmer(std_length ~ age + temp_life + age * method  + (1|season), data = sander1, REML = F)
P2s <- lmer(std_length ~ age * temp_life + age * method + (1|season), data = sander1, REML = F)
anova(P1, P2s)

P3 <- lmer(std_length ~ fage + temp_life + fage * method  + (1|season), data = sander1, REML = F)
P4s <- lmer(std_length ~ fage * temp_life + fage * method + (1|season), data = sander1, REML = F)
anova(P3, P4s)

```

```{r}
plot_model(P2s,
                   type = "pred", 
                   terms = c("temp_life", "age"),
                   show.data = TRUE,
                   pred.type = c("fe"),
                   title = "",
                   show.legend = TRUE,
                   jitter = 0.1,
                   axis.title = c("Average fish life temperature", "Standard length (cm)"),
                   show.values = FALSE) +
  theme_bw() +  # Set plot theme to black and white
  theme(panel.background = element_rect(fill = "white"),  # Set background color to white
        axis.text = element_text(size = 14),  # Increase size of axis text
        axis.title = element_text(size = 16),  # Increase size of axis titles
        panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank())  # Remove minor grid lines

```

```{r}
# Summarise neatly in a table
tab_model(P2s,
          show.zeroinf = F,
             dv.labels = c("Pikeperch"),
           string.pred = "Coefficient",
             string.ci = "Conf. Int (95%)",
              string.p = "P-value",
               p.style = c("numeric"),
                emph.p = FALSE,
             transform = NULL)
```

Using age as factor. Model with different slopes for different age group.
```{r}
plot_model(P4s,
                   type = "pred", 
                   terms = c("temp_life", "fage"),
                   show.data = TRUE,
                   pred.type = c("fe"),
                   title = "",
                   show.legend = TRUE,
                   jitter = 0.1,
                   axis.title = c("Average fish life temperature", "Standard length (cm)"),
                   show.values = FALSE) +
  theme_bw() +  # Set plot theme to black and white
  theme(panel.background = element_rect(fill = "white"),  # Set background color to white
        axis.text = element_text(size = 14),  # Increase size of axis text
        axis.title = element_text(size = 16),  # Increase size of axis titles
        panel.grid.major = element_blank(),  # Remove major grid lines
        panel.grid.minor = element_blank())  # Remove minor grid lines

```


Residuals
```{r}
plot(P2s)
resid <- residuals(P2s)
hist(resid, breaks = 30, main = "Histogram of Residuals")
qqnorm(resid)
qqline(resid, col = "red")

```

```{r}
# Influential observations in the model? 
# Measure using Cook's distance
par(mfrow = c(1, 1))
plot(cooks.distance(P2s), 
     type = "h",
     xlab = "Observation", 
     ylab = "Cook's distance",
     ylim = c(-0.1, 1.2))
abline(h=1, lty = 2)
abline(h=0, lty = 2)
# All below 1 - that is fine
```


# One plot


```{r}
library(cowplot)

create_plot <- function(model, species_name, show_x_axis_text = TRUE) {
  plot <- plot_model(model, type = "pred", terms = c("temp_life", "age"), show.data = TRUE, pred.type = "fe") +
    ylim(0, NA) +
    scale_y_continuous(limits = c(0, NA), breaks = c(0, 10, 20, 30, 40, 50, 60, 70)) +
    scale_x_continuous(limits = c(8, 10.5), breaks = c(8, 9, 10)) +  # Force x-axis limits and breaks
    labs(x = NULL, y = NULL) +  # Remove x and y axis labels from individual plots
    ggtitle(species_name) +
    theme_bw() +
    theme(
      panel.background = element_rect(fill = "white"),
      axis.text = element_text(size = 17),
      axis.title = element_text(size = 17),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title = element_text(size = 17, hjust = 0.5),
      legend.text = element_text(size = 16),  
      legend.title = element_text(size = 16)
    )

  # Conditionally remove x-axis text but keep the ticks
  if (!show_x_axis_text) {
    plot <- plot + theme(axis.text.x = element_blank())  # Hide only the axis text
  }
  
  return(plot)
}


# Species names
species_names <- c("Bream", "Roach", "Perch", "Pikeperch")

# Create individual plots with species names
# Remove x-axis text (but keep ticks) for the upper plots
plot1 <- create_plot(P2b, species_names[1], show_x_axis_text = FALSE)
plot2 <- create_plot(P2r, species_names[2], show_x_axis_text = FALSE)

# Keep x-axis text and ticks for the bottom plots
plot3 <- create_plot(P2p, species_names[3], show_x_axis_text = TRUE)
plot4 <- create_plot(P2s, species_names[4], show_x_axis_text = TRUE)

# Combine plots using cowplot
combined_plot <- plot_grid(
  plot1, plot2, plot3, plot4, 
  label_size = 17, 
  ncol = 2
)

# Adjust the positions of the labels to avoid overlap
final_plot <- ggdraw() +
  draw_plot(combined_plot, 0.02, 0.05, 1, 0.95) +  # Shift the combined plot up
  draw_label("Length (cm)", x = -0.01, angle = 90, vjust = 1, hjust = 0.5, size = 17) +
  draw_label("Average fish life temperature (C°)", y = -0.02, vjust = -1, hjust = 0.5, size = 17) + 
  theme(plot.margin = margin(15, 15, 15, 15))

# Print the final plot
print(final_plot)

ggsave("final_plot_x3.tiff", plot = final_plot, units = "mm", 
       width = 300, height = 250, dpi = 300, compression = 'lzw')
```


Plot showing results from different aging methods

```{r}
# Function to create a plot for a given model
create_plot <- function(model, species_name, show_x_axis_text = TRUE) {
  plot <- plot_model(model, type = "pred", terms = c("temp_life", "age", "method"), show.data = TRUE, pred.type = "fe") +
    ylim(0, NA) +
    labs(x = NULL, y = NULL) +  # Remove x and y axis labels from individual plots
    ggtitle(species_name) +
    theme_bw() +
    theme(
      panel.background = element_rect(fill = "white"),
      axis.text = element_text(size = 14),
      axis.title = element_text(size = 16),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.title = element_text(size = 18, hjust = 0.5)
    )

  # Conditionally remove x-axis text
  if (!show_x_axis_text) {
    plot <- plot + theme(axis.text.x = element_blank())
  }
  
  return(plot)
}

# Species names
species_names <- c("Bream", "Perch", "Pikeperch", "Roach")

# Create individual plots with species names
plot1 <- create_plot(P2b, species_names[1], show_x_axis_text = TRUE)
plot2 <- create_plot(P2p, species_names[2], show_x_axis_text = TRUE)
plot3 <- create_plot(P2s, species_names[3], show_x_axis_text = TRUE)
plot4 <- create_plot(P2r, species_names[4], show_x_axis_text = TRUE)

# Combine plots using cowplot
combined_plot <- plot_grid(
  plot1, plot2, plot3, plot4, 
  label_size = 16, 
  ncol = 2
)

# Adjust the positions of the labels to avoid overlap
final_plot <- ggdraw() +
  draw_plot(combined_plot, 0.02, 0.05, 1, 0.95) +  # Shift the combined plot up
  draw_label("Length (cm)", x = -0.01, angle = 90, vjust = 1, hjust = 0.5, size = 16) +
  draw_label("Average fish life temperature (C°)", y = -0.02, vjust = -1, hjust = 0.5, size = 16) + theme(plot.margin = margin(15, 15, 15, 15))

print(final_plot)

```
